---
title: "Quantile Regression"
author: "Iris Zhang"
date: "2023-12-17"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



https://drkebede.medium.com/quantile-regression-tutorial-in-r-f2eec72c132b 



```{r}
# packages
library(tidyr)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(quantreg)
library(mice)
```



## load data (cleaned and imputed dataset)


```{r}
# data dictionary
load("data/nhanes_key.rda")
load("data/key_guide.rda")


```



## select covariates 

1. `bp_sys_mean`: Systolic blood pressure (SBP), mm Hg
2. `bp_dia_mean`: Diastolic blood pressure (DBP), mm Hg
3. `demo_age_years`: Age, years
4. `demo_gender`: Gender 
5. `cc_bmi`: Body mass index, kg/m2
6. `chol_total`: Total cholesterol, mg/dL
7. `chol_hdl`: HDL cholesterol, mg/dL
8. `chol_ldl`: LDL cholesterol, mg/dL
9. `INDFMPIR`: Family PIR 
10. `DMDEDUC2`: Education level - Adults 20+
11. `DMDMARTL`: Marital status
12. `ALQ130`: Avg # alcoholic drinks/day - past 12 mos
13. `SMQ020`: Smoked at least 100 cigarettes in life
 








#### Only Keep rows have BP values 


```{r}
# Only dropping NAs in bp_sys_mean results in 56534 
# Dropping NAs in both bp_sys_mean and bp_dia_mean results in 56286 
dataset.bp <- dataset %>% drop_na(bp_sys_mean, bp_dia_mean)
```



```{r}
# Dropping NAs in cc_bmi results in 55,121 
# Dropping NAs in chol_total results in 55,121 
# Dropping NAs in INDFMPIR results in 51,136
# Dropping NAs in DMDEDUC2 results in 52,384
# Dropping NAs in DMDMARTL results in 54,010
# Dropping NAs in ALQ130 results in 33,216
# Dropping NAs in SMQ020 results in 53,336 
# Dropping NAs in chol_hdl results in 24,043
# Dropping NAs in chol_ldl results in 24,020 


# Dropping NAs in all above results in 12,892 
# Dropping NAs in cc_bmi,chol_total,INDFMPIR,DMDEDUC2,DMDMARTL,ALQ130,and SMQ020 results in 12,903 

dataset.test <- dataset.bp %>% drop_na(cc_bmi,chol_total,INDFMPIR,DMDEDUC2,DMDMARTL,ALQ130,SMQ020,chol_hdl,chol_ldl)
```



### Check distributions of survey year 

```{r}
ggplot(dataset.test, aes(x=svy_year)) + geom_bar() +  labs(x='svy_year')
```


### Some EDA plots 


```{r}
ggplot(dataset.test, aes(x=demo_age_years, 
                          y=bp_sys_mean, col=cc_bmi)) + geom_point() +
geom_smooth(method = "lm", col="blue") +
      geom_quantile(color = "red", quantiles = 0.99)+
  xlab("Age") + ylab("Systolic blood pressure")

```









## Quantile Analysis 


```{r}
rqfit <- rq(bp_sys_mean ~ demo_age_years+cc_bmi+chol_total, tau=0.90,data = dataset.test)
summary(rqfit)
```










